---
title: "Models"
execute: 
  echo: false
  message: false
  warning: false
---

```{r}
library(dplyr)
library(readr)
library(remotes)
library(primer.data)
library(tidyverse)
x <- read_csv("data/diabetes_prediction_dataset.csv")
y <- read_csv("data/diabetes.csv")
```

 

## Age and Gender

```{r}
ag_x <- x |>
  mutate(diabetes = replace(diabetes, diabetes == 0, NA)) |>
  drop_na()|>
  select(gender, age, diabetes)  
ig_y <- y |>
  drop_na() |>
  select(Glucose, Insulin, Outcome)
bmi_y <- y |>
    drop_na() |>
    select(BMI, Outcome)
bp_y <- y |>
  drop_na() |>
  select(BloodPressure, Outcome)
```

```{r}
ag_x |>
  ggplot(aes(x = age, fill = gender)) +
  geom_histogram(bins = 25) +
  scale_fill_manual(values = c("#bfabd6", "#77b2ed")) +
  labs(title = "Age and Gender Distribution of People with Diabetes", 
       x = "Age", 
       y = "", 
       fill = "Gender", 
       caption = "Source: Kaggle (Mohammed Mustafa)") +
  theme_classic()
```

**Take-away:** It is more common for females to have diabetes than males. As age increases, the likeliness of getting diabetes becomes higher and higher. There is a spike at around 80 years old with almost 14 percent of the sample having the disease.

 

## Factors 1 & 2

```{r}
ig_y |>
  ggplot(aes(x = Glucose, y = Insulin, color = factor(Outcome))) +
  geom_smooth(method = "loess", formula = y ~ x, se = FALSE) +
  scale_color_manual(values = c("#bfabd6", "#77b2ed"), 
                     labels = c("No Diabetes", "Diabetes")) +
  labs(title = "Insulin and Glucose Levels of People with and without Diabetes", 
       x = "Glucose Level", 
       y = "Insulin Level", 
       color = "Diabetes?", 
       caption = "Source: Kaggle (Akshay Dattatray Khare)") +
theme(
  axis.title.x = element_text(face = "bold"),
  axis.title.y = element_text(face = "bold"),
  axis.text.x = element_text(face = "bold"),
  axis.text.y = element_text(face = "bold"),
  legend.title = element_text(face = "bold"),
  panel.grid = element_blank())

```

Among individuals with glucose levels below 100 in this sample, diabetic patients have higher insulin levels compared to non-diabetic patients. However, for individuals with glucose levels exceeding 100, diabetic patients show lower insulin levels in comparison to non-diabetic patients, indicating a potential impairment in glucose regulation among individuals with diabetes.

 

## Factor 3

```{r}
bmi_y |>
  ggplot(aes(x = Outcome, y = BMI, fill = factor(Outcome))) +
  geom_boxplot() + 
  scale_fill_manual(values = c("#bfabd6", "#77b2ed"),
                    labels = c("No Diabetes", "Diabetes")) +
  labs(title = "BMI Levels for People with and without Diabetes",
       x = "",
       y = "BMI",
       fill = "Diabetes?", 
       caption = "Source: Kaggle (Akshay Dattatray Khare)") +
  theme_classic() +
  theme(plot.title = element_text(face = "bold", size = 14),
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank())
```

The higher position of the boxplot for the diabetes group suggests a correlation between diabetes and higher BMI levels as there would be a disruption in glucose metabolism.

 

## Factor 4

```{r}
bp_y |>
  ggplot(aes(x = Outcome, y = BloodPressure, fill = factor(Outcome))) +
  geom_boxplot() +
  scale_fill_manual(values = c("#bfabd6", "#77b2ed"),
                    labels = c("No Diabetes", "Diabetes")) +
  labs(title = "Blood Pressure Levels for People with and without Diabetes",
       x = "",
       y = "Blood Pressure",
       fill = "Diabetes?", 
       caption = "Source: Kaggle (Akshay Dattatray Khare)") +
  theme_classic() +
  theme(plot.title = element_text(face = "bold", size = 14),
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank())
```

The boxplot for the diabetes group is only slightly higher than that of the non-diabetes group which suggests a weak correlation between diabetes and blood pressure levels. There could be other factors that play a more significant role in influencing blood pressure outcomes.
