---
title: "Models"
execute: 
  echo: false
  message: false
  warning: false
---

```{r}
library(dplyr)
library(readr)
library(remotes)
library(primer.data)
library(tidyverse)
library(rstanarm)
x <- read_csv("data/diabetes_prediction_dataset.csv")
y <- read_csv("data/diabetes.csv")
```

 

## Age and Gender

```{r}
ag_x <- x |>
  mutate(diabetes = replace(diabetes, diabetes == 0, NA)) |>
  drop_na()|>
  select(gender, age, diabetes)
ig_y <- y |>
  drop_na() |>
  select(Glucose, Insulin, Outcome)
bmi_y <- y |>
    drop_na() |>
    select(BMI, Outcome)
bp_y <- y |>
  drop_na() |>
  select(BloodPressure, Outcome)
```

```{r}

ag_x |>
  ggplot(aes(x = age, fill = gender)) +
  geom_histogram(bins = 25) +
  scale_fill_manual(values = c("#77b2ed", "#bfabd6")) + 
  labs(title = "Age and Gender Distribution of People with Diabetes", 
       x = "Age", 
       y = "", 
       fill = "Gender", 
       caption = "Source: Kaggle (Mohammed Mustafa)") +
  theme_classic()

```

**Take-away:** It is more common for males to have diabetes than females. As age increases, the likeliness of getting diabetes becomes higher and higher. There is a spike at around 80 years old with almost 14 percent of the sample having the disease.

 

## Data Model

```{r}
poffit <- x |> 
  drop_na() |>
  select(age, gender, diabetes) |> 
  slice_sample(n = 1250)
model <- stan_glm(data = poffit, 
                  formula = diabetes ~ age + gender, 
                  family = binomial, 
                  refresh = 0,
                  seed = 9)
print(model, digits = 2)
```

**Analysis:** The age and genderMale median being positive shows that males are more likely to have diabetes than females and that the likeliness of diabetes increases with age. However, the median absolute deviation indicates that there is some uncertainty regarding the association between gender, age, and diabetes.
